---
title: "Prediction Assignment - Practical Machine Learning"
author: "Coursera_S"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this project, we use accelerometer data to predict the manner in which participants performed barbell lifts. The variable `classe` is the target.

## Load Libraries

```{r}
packages <- c("caret", "randomForest", "dplyr", "e1071")
install_if_missing <- function(p) if (!require(p, character.only = TRUE)) install.packages(p, dependencies = TRUE)
invisible(lapply(packages, install_if_missing))

library(caret)
library(randomForest)
library(dplyr)
library(e1071)
```

## Load and Clean Data

```{r}
training <- read.csv("pml-training.csv", na.strings = c("NA", "", "#DIV/0!"))
training <- training[, colSums(is.na(training)) == 0]
training <- training %>%
  select(-c(user_name, raw_timestamp_part_1, raw_timestamp_part_2, 
            cvtd_timestamp, new_window, num_window, X))
training$classe <- as.factor(training$classe)
```

## Data Partitioning

```{r}
set.seed(1234)
inTrain <- createDataPartition(training$classe, p = 0.7, list = FALSE)
trainData <- training[inTrain, ]
testData  <- training[-inTrain, ]
```

## Model Training

```{r}
set.seed(1234)
control <- trainControl(method = "cv", number = 5)
modelFit <- train(classe ~ ., data = trainData, method = "rf", trControl = control)
```

## Model Evaluation

```{r}
predictions <- predict(modelFit, newdata = testData)
confusionMatrix(predictions, testData$classe)
```

## Final Predictions (Optional)

```{r eval=FALSE}
testing <- read.csv("pml-training.csv", na.strings = c("NA", "", "#DIV/0!"))
testing <- testing[, colSums(is.na(testing)) == 0]
testing <- testing %>%
  select(-c(user_name, raw_timestamp_part_1, raw_timestamp_part_2, 
            cvtd_timestamp, new_window, num_window, X))
final_predictions <- predict(modelFit, newdata = testing)
final_predictions
```

## Conclusion

We trained a Random Forest model using 5-fold cross-validation. The model shows high accuracy in predicting the manner in which exercises were performed. The `classe` variable is predicted with high confidence.
